A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-21 text/manuscript.Rmd

It does not matter for our purposes whether a single tumor is “enriched” for
multiple pathway mutations, chiefly because such metrics favor larger pathways (
_i.e._ pathways with more member genes) over smaller ones, and non-expressed
genes over expressed ones due to transcription coupled repair
[@kandothMutationalLandscapeSignificance2013; @kimPancancerAnalysisSomatic2018;
@Pervasive_lesion_segregation]. Data selected in this manner are likely to be
noisy for several reasons. First, point mutations can be deleterious
(attenuating or hypomorphic) or activating (neomorphic) in genes that act
functionally as either activators or repressors. Within the scope of this study
we must accept this uncertainty and assume mutations are generically disruptive
to the cell regardless of whether they are activating or deactivating or whether
they affect a gene that we might classify as one thing or another ( _e.g._
"Tumor suppressor" or "Oncogene"). Second, as previously mentioned, we know that
low-expressed genes and non-expressed genes are known to accumulate mutations at
a higher rate due to transcription coupled repair. To address this issue, we
identified genes with low expression in each type of cancer and eliminated them
for that cancer type only (see methods for details). <strike>Highly expressed
genes also have high mutation rates owing to transcription induced mutagenesis
[@parkGenomicEvidenceElevated2012]. We felt that this mechanism results in
cell-type-specific biases that might be biologically meaningful for
predisposition to different classes of cancer however, and therefore chose to
include these genes in our analysis.</strike>

After selecting our pathways and genes, we then compiled a matrix of the
pathways, assigning a Boolean value (a 1 or 0, indicating whether any pathway
member gene carries a mutation) to the pathway regardless of the number of genes
carrying mutations (this data is visualized in **figure 1a**).

>>>>

It does not matter for our purposes whether a single tumor is “enriched” for
multiple pathway mutations, chiefly because such metrics favor larger pathways (
_i.e._ pathways with more member genes) over smaller ones, and non-expressed
genes over expressed ones due to transcription coupled repair
[@kandothMutationalLandscapeSignificance2013; @kimPancancerAnalysisSomatic2018;
@Pervasive_lesion_segregation; @expression-aware-annotation]. Data selected in
this manner are likely to be noisy for several reasons. First, point mutations
can be deleterious (attenuating or hypomorphic) or activating (neomorphic) in
genes that act functionally as either activators or repressors. Within the scope
of this study we must accept this uncertainty and assume mutations are
generically disruptive to the cell regardless of whether they are activating or
deactivating or whether they affect a gene that we might classify as one thing
or another ( _e.g._ "Tumor suppressor" or "Oncogene"). Second, as previously
noted, we know that low-expressed genes and non-expressed genes are known to
accumulate mutations at a higher rate due to transcription coupled repair. To
address this issue, we identified genes with low expression in each type of
cancer and eliminated them for that cancer type only (see methods for details).
<strike>Highly expressed genes also may have high mutation rates owing to
transcription induced mutagenesis [@parkGenomicEvidenceElevated2012]. We felt
that this mechanism results in cell-type-specific biases that might be
biologically meaningful for predisposition to different classes of cancer
however, and therefore chose to include these genes in our analysis.</strike>
After selecting our pathways and genes, we then compiled a matrix of the
pathways, assigning a Boolean value (a 1 or 0, indicating whether any pathway
member gene carries a mutation) to the pathway regardless of the number of genes
carrying mutations (this data is visualized in **figure 1a**).
